---
title: "Pipeline for development"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pipeline for development}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

**Rationale**

This document will serve as a template for the development of the pipeline. We can run existing function inside a code chunk. Custom code (which may eventually become a function) can be written inside a chunk with comments. Eventually this pipeline will call the R package itself.

-   Run functions in chunks

-   Run custom code with comments

-   Produce record of analysis

-   Embed in line text

-   Embed figures

# Setup

## Project setup

To run this pipeline locally, you need to setup a R project in Rstudio by selecting `File` and `New project` checkout the repo <https://github.com/PARADISE-AAV/rkd_package/> using git.

Then install the package using **Ctrl** + **Shift** + **B** or equivalent.

### Data protection warning

Please familiarise yourself with github before using this pipeline. **Do not store data files locally within the folder as these may be pushed to github.** Link to the onedrive location of the files on your computer.

## load libraries 

We can make all functions of the package available by running

```{r library, include=FALSE}
library(rkdpipeline)
message(packageVersion('rkdpipeline'))
```

## Paths to data

in the future we may have a static path to a onedrive url or an API to Repcap. For now the paths should lead to a local sync of one drive folders.

There are 3 types of data

-   RKD dataset
-   Biomarker master
-   FreezerWorks exports

For this pipeline to work each user should set these paths each time.
Note the following code is not reproducible for other users.

```{r paths}

Biomarker_path <- "C:/Users/COQM/OneDrive - Trinity College Dublin/Biomarker_master/Biomarker master_Data only.csv"
   
fw_path <- "C:/Users/COQM/OneDrive - Trinity College Dublin/Freezerworks exports/Freezerworks Exports"

#specific RKD version
rkd_file <- "C:/Users/COQM/OneDrive - Trinity College Dublin/Redcap Data exports/RareKidneyDiseaseRIT-PARADISE_DATA_LABELS_2025-07-14_2214.csv"
```

# Pipeline

## Import the data

This function will load the RKD dataset from disk.
For the purposes of this example, we will pull in whichever CSV file is in the current working directory.

```{r load_RKD}
rkd <- load_riv(rkd_file)
colnames(rkd)[1:10]
```

## Clean the data

The CleanRKD function clean the in 5 steps:

* set the date variable to the format `"%Y-%m-%d"`
-   reduce the ethnicity to 6 groups where different subgroup are regrouped
-   clean some `RKD.ID` trouble to be sure that we have no problem when we do any merge with other dataset
-   select the variable present only for the Encounters data and the Initial data (demographics, diagnostics and all other exams performed at the moment of the diagnostics) and merge the Encounter data and Initial data that the initial data are replicate for each Encounter
-   create the following variable Age of the Encounter and ANCA titration (Anti MPO or Anti PR3 or NA)

```{r clean_rkd)}
output_dir <- tempdir()
split_RKD <- SplitRIV(rkd, tempdir())
names(split_RKD)
```




## Use the Demographic filter

The DemographicFilterRKD function use the Paradise filter:

* SVV = GPA / MPA
-   AND diagnosis confidence = definite
-   AND SVV (IC) != Anti-GBM, IgA, cryo
-   AND Secondary vasculitis != Yes
-   AND 'Other' != Yes
-   AND 'Medium vessel' != 1 or 2
-   AND 'large vessel' != 1 or 2
-   AND 'variable vessel' != 1 or 2
-   AND (ANCA subtype = PR3 / MPO / both) OR (biopsy = Definite vasculitis AND histologically confirmed = YES)


```{r demographic filter}
Inclusion_Criteria_rkd <- FilterRIV(split_RKD$Initial, tempdir(),c("Definite MPA", "Definite GPA"))
head(Inclusion_Criteria_rkd)[, 1:10]
```

# Clean General characteristics data

```{r}

clean_rkd <- clean_riv(Inclusion_Criteria_rkd,tempdir())

```
```{r}

ANCAimputation=CPD_ANCA_Imputation(split_RKD$Encounter,tempdir())
ISmedImputation=CPD_IS_medication_Imputation(ANCAimputation,tempdir())
ISstatusImputation=CPD_IS_status_Imputation(ISmedImputation,tempdir())
clean_encounter=CPD_Encounter(ISstatusImputation, tempdir())

```


# merge Encounters data and General Characteristics data

```{r}

merge_rkd <- Merge_Encounter_initial(clean_rkd, clean_encounter, tempdir())

```

```{r}
IVTherapy = CPD_IVTherapy_interval(split_RKD, tempdir())
```

```{r}
Medication = CPD_Continuous_Medication_interval(split_RKD, tempdir())
```

```{r}
IVTherapy_treatment=CPD_IVTherapy_Treatment(IVTherapy, merge_rkd, tempdir())
```

```{r}
Medication_treatment=CPD_Medication_Treatment(Medication, merge_rkd, tempdir())
```

```{r}
CPDRenal=CPD_Renal(split_RKD,tempdir())
```



```{r classify jrelapse}
Classify_rkd_All <- ClassifyRIVEncounter(merge_rkd, tempdir(), "All", CM_data=Medication_treatment, IV_data=IVTherapy_treatment, Renal=CPDRenal)
head(Classify_rkd_All)[, 1:10]

```


# Appendix

```{r sessioin}
sessionInfo()
```