---
title: "Tidy data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tidy data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r preamble}
library(rivpipeline)
input_filename <- list.files(pattern = 'PARADISE_DATA_2.*\\.csv')
input_filename
```

Import the file described above.
There are two types of files, `DATA`, which has unambiguous column names and categorical values, and `DATA_LABELS` with the human-friendly labels.
In Redcap somewhere there should be a mapping between `DATA` and `LABELS`.

```{r}
library(readr)
redcap_df <- read_csv(input_filename, guess_max = Inf, show_col_types = FALSE)
redcap_df[1:6, ]
```

What data types do we have here?

```{r}
unique(redcap_df$redcap_repeat_instrument)
```

Split the data into lots of different tables.
This is what makes the data tidy---having separate tables for different kinds of measurements.

```{r}
library(tidyr)
nested_data <- redcap_df %>% nest(.by = redcap_repeat_instrument)
nested_data
```

Now we can strip out columns that are all NAs and get a list of columns associated with each data type.

```{r, message=FALSE}
library(tidyverse)

drop_na_cols <- function(df) {
  stopifnot(is.data.frame(df))
  # Drop columns that are all NA
  df <- df[(colSums(!is.na(df)) > 0)]
  # Drop columns that are all constant
  # df <- df[map_lgl(df, \(x) length(unique(x)) > 1)]
  df
}

nested_data %>%
  mutate(cleaned = map(data, drop_na_cols),
         cols_orig = map_int(data, ncol),
         cols_clean = map_int(cleaned, ncol)) %>%
  add_count(wt = cols_clean, name = 'n_clean_total')
```

What are the common columns?

```{r}
# Common to add data types
nested_data %>%
  mutate(cleaned = map(data, drop_na_cols),
         clean_cols = map(cleaned, colnames)) %>%
  pull(clean_cols) %>%
  reduce(intersect)

# Common to two or more data types
nested_data %>%
  mutate(cleaned = map(data, drop_na_cols),
         clean_cols = map(cleaned, colnames)) %>%
  select(clean_cols) %>%
  unnest_longer(clean_cols) %>%
  count(clean_cols) %>%
  filter(n > 1)
```



### Advanced pivoting

See <https://tidyr.tidyverse.org/reference/pivot_longer_spec.html>

```{r}
# coltypes <- sapply(redcap_df, class)
# redcap_df %>%
#   build_longer_spec(-(record_id:patient_id),
#                     names_to = 'variable',
#                     values_to = 'value')
```


## Clean RIV 

In this section we demonstrate how to recreate the `clean_riv` function using the new, cleaner data structure.

Turn the table into a list of separate tables:

```{r}
db <- nested_data %>%
  mutate(cleaned = map(data, drop_na_cols),
         clean_cols = map(cleaned, colnames)) %>%
  select(redcap_repeat_instrument, cleaned) %>%
  # Rename the baseline table from `NA` to 'baseline'
  mutate(redcap_repeat_instrument = replace_na(redcap_repeat_instrument, 'baseline')) %>%
  # Convert to a named list of tables
  tibble::deframe()

str(db, max.level = 1)
```

*Advanced*: We could access this list like a database, but might be easier to assign them in an [environment](https://adv-r.hadley.nz/environments.html).
If we assign them to the `.GlobalEnv` then we can manipulate the tables just like any R object.

```{r, eval = FALSE}
list2env(db, envir = .GlobalEnv) # creates an object in R for each table
```

However, it's easier to keep track of a list and this should be more intuitive to less advanced R users.
When we save the results it can then be individual tables to CSV or the whole list to an `rds` or `RData` file.

The `clean_riv` function performs the following steps:

1. parse dates
2. collapse ethnicity
3. fix IDs
4. tidy encounters
5. compute patient's age at each encounter
6. Anti MPO PR3(?)
7. compute interval between diagnosis and date of visit
8. save to file

### Parse dates

This can be mostly skipped since we imported the data using `readr::read_csv`, which automatically converts date-like columns to dates.
There is only one column containing the name "date_of_" that isn't already in `Date` format.

```{r}
col_types <- map_chr(redcap_df, class)
date_cols <- str_subset(colnames(redcap_df), 'date')
which(col_types[date_cols] != 'Date')
```

This lives in the baseline table:

```{r}
names(db)[map_lgl(db, ~ 'date_of_symptons_choice' %in% names(.x))]
unique(db$baseline$date_of_symptons_choice)
```
Clearly this is a binary categorical variable so doesn't need parsing either.

### Collapse ethnicity

Ethnicity is defined at baseline and doesn't change with encounters.

```{r}
unique(db$baseline$ethnicity)
```

Because we are not using the `LABELS` version of the dataset, this is already numerically coded.
See <https://3.basecamp.com/3790396/buckets/4192102/uploads/4462366541> for the mapping of field names to labels and field attributes.

```{r}
# Better to keep this in a dictionary file somewhere, but just shown for demonstration.
ethnicity_codes <- tribble(
  ~code, ~ethnicity_label,
  15, 'W2 - White Irish',
  14, 'W1 - White British',
  16, 'W9 - Any other White background',
  1, 'A1 - Indian',
  2, 'A2 - Pakistani',
  3, 'A3 - Bangladeshi',
  4, 'A9 - Any other Asian background',
  5, 'B1 - Caribbean',
  6, 'B2 - African',
  7, 'B9 - Any other Black background',
  8, 'M1 - White and Black Caribbean',
  9, 'M2 - White and Black African',
  10, 'M3 - White and Asian',
  11, 'M9 - Any other Mixed background',
  12, 'O1 - Chinese',
  13, 'O9 - Any other ethnic group',
  17, 'NS - Not Stated'
)

# Convert the first letter to simplified category labels
ethnicity_map <- c(
  W = 'White',
  A = 'Asian',
  B = 'Black',
  M = 'Mixed ethnicity',
  O = 'Other',
  N = 'Not Stated'
)

db$baseline <- left_join(db$baseline, ethnicity_codes, by = c(ethnicity = 'code')) %>%
  mutate(ethnicity = ethnicity_map[str_sub(ethnicity_label, 1, 1)])

# Verify output
unique(db$baseline$ethnicity)
```

### Fix IDs

In the non-`LABELS` file, there are no IDs (here called `record_id` but in the `LABELS` file called `RKD ID`) equal to `-` so the `rkd_fix_ids` function is unnecessary.

```{r}
sum(is.na(redcap_df$record_id))
```

So there is nothing to be done.

### Tidy encounters

The `rkd_tidy_encounters` function is probably where the advantage of separate tables will be most keenly felt.
The inside of that function is essentially making two tables: `rkd_initial`, corresponding to our "baseline" table (i.e. `redcap_repeat_instrument` missing), and `rkd_encounters` corresponding to longitudinal measurements.

Here we already have the table `baseline` and the various longitudinal tables and at any time we can join them together on the `record_id` (RKD ID).

### Compute patient's age

This is the first time that we need to join one table with another.
Specifically, the `date_of_visit` lives in `encounters` whereas `date_of_birth` is a `baseline` variable.

```{r}
dob <- db$baseline %>% select(record_id, date_of_birth)

db$encounters <- db$encounters %>%
  left_join(dob, by = 'record_id') %>%
  mutate(age_at_encounter = interval(date_of_birth, date_of_visit) / years(1))

# Examples
db$encounters %>%
  select(date_of_visit, date_of_birth, age_at_encounter) %>%
  sample_n(6)
```

### Anti MPO PR3(?)

This is just a very simple data cleaning operation.
Indicate which out of Anti-MPO and Anti-PR3 level is present.
If neither, return `NA` and if both, return whichever is higher.

Let's check which table this lives in...

```{r}
names(db)[map_lgl(db, ~ 'anti_mpo_level' %in% colnames(.x))]
```

OK now let's define the variable.

```{r}
db$encounters <- db$encounter %>%
  mutate(anti_mpo_pr3 = case_when(
    is.na(anti_mpo_level) & is.na(anti_pr3_level) ~ NA,
    is.na(anti_mpo_level) & !is.na(anti_pr3_level) ~ "PR3",
    !is.na(anti_mpo_level) & is.na(anti_pr3_level) ~ "MPO",
    anti_mpo_level > anti_pr3_level ~ "MPO",
    TRUE ~ "PR3"
  )
)

db$encounters %>%
  select(anti_mpo_pr3, anti_mpo_level, anti_pr3_level) %>%
  sample_n(5)
```

### Compute interval between diagnosis and date of visit

Same idea as calculating patient age (above).
Left as an exercise to the reader...

### Save to file

Either save the whole list or write out each frame individually.

```{r, eval = FALSE}
saveRDS(db, 'riv_database.rds')
# or
for (i in names(db)) {
  write_csv(db[[i]], paste0(names(db)[i], '.csv'))
}
```

## Medications

Start and stop dates of medications.

```{r}
names(db)[map_lgl(db, ~ any(c('start_date', 'stop_date') %in% colnames(.x)))]
```