---
title: "Pipeline_for_development"
author: "Conor Finlay/Matthieu Coq"
date: '`r Sys.Date()`'
output: html_document
---

**Rationale**

This document will serve as a template for the development of the pipeline. We can run existing function inside a code chunk. Custom code (which may eventually become a function) can be written inside a chunk with comments. Eventually this pipeline will call the R package itself.

-   Run functions in chunks

-   Run custom code with comments

-   Produce record of analysis

-   Embed in line text

-   Embed figures

# Setup

Load required libraries and set markdown settings (for plotting, caching, etc)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project setup

To run this pipeline locally, you need to setup a R project in Rstudio by selecting `File` and `New project` checkout the repo <https://github.com/PARADISE-AAV/rkd_package/> using git.

### Data protection warning

Please familiarise yourself with github before using this pipeline. **Do not store data files locally within the folder as these may be pushed to github.** Link to the onedrive location of the files on your computer.

## load libraries 

```{r libary, include=FALSE}
library(tidyverse)
```

## Working directories 

The `.Rproj` file (which is ignored by GitHub and specific for each local instance of the repo) sets the working directory to wherever the project is located on your computer. This has the advantage of simplifying pathing by making them relative to this folder.

```{r pathing}
#print local directory 
getwd()
list.files()
```

## Paths to data

in the future we may have a static path to a onedrive url or an API to Repcap. For now the paths should lead to a local sync of one drive folders.

There are 3 types of data

-   RKD dataset

-   Bio marker Master

-   freezer works exports

For this pipeline to work each user should set these paths each time.

```{r paths}
#paths
RKD_path <- "C:/Users/Administrator/OneDrive - Trinity College Dublin/Redcap Data exports/"

Biomarker_path <- "C:/Users/Administrator/Trinity College Dublin/THKC - Documents/Arlena Carney/Makala Carroll/Projects/Mini Biobank/Biomarker_master/Biomarker master-Data only.csv"
   
fw_path <- "C:/Users/Administrator/OneDrive - Trinity College Dublin/Freezerworks exports"

#specific RKD version
RKD_file <- "C:/Users/Administrator/OneDrive - Trinity College Dublin/Redcap Data exports/RareKidneyDiseaseRIT-FullVasculitisExport_DATA_LABELS_2023-04-24_1444.csv"
```

## Functions

Here we will load the approved functions into the R session. It assumes that all functions are required.

```{r load_functions}
# find all files ending in ".R" and run source
setwd("~/GitHub/rkd_package/function")
for (f in list.files(pattern = "*.R$")) {
    source(f)
}

```

# Pipeline

## loadRKD
This function will load RKD dataset
```{r load_RKD}
setwd("~/GitHub/rkd_package")
rkd  <- LoadRKD(files_name = RKD_file)
head(rkd)[1:5,1:10]
```


```{r wd_check, warning=FALSE}
a=getwd()
dir.create(paste(a,"/output",sep=""))
```
## Clean the data

The CleanRKD function clean the in 5 step:

* set the date variable to the format "%Y-%m-%d"
-   reduce the ethnicity to 6 group where different subgroup are regrouped
-   clean some RKD.ID trouble to be sure that we have no problem when we do any merge with other dataset
-   select the variable present only for the Encounters data and the Initial data (demographics, diagnostics and all other exams performed at the moment of the diagnostics)and merge the Encounter dtaa and Initail data that the initial data are replicate for each Encounter
-   create the following variable Age of the Encounter and ANCA titration (Anti MPO or Anti PR3 or NA)

```{r, warning=FALSE}

clean_rkd <- CleanRKD(rkd, paste(a,"/output",sep=""))
head(clean_rkd)[1:5,1:10]
```
## Use the Demographic filter

The DemographicFilterRKD function use the Paradise filter:

* SVV = GPA / MPA
-   AND diagnosis confidence = definite
-   AND SVV (IC) != Anti-GBM, IgA, cryo
-   AND Secondary vasculitis != Yes
-   AND 'Other' != Yes
-   AND 'Medium vessel' != 1 or 2
-   AND 'large vessel' != 1 or 2
-   AND 'variable vessel' != 1 or 2
-   AND (ANCA subtype = PR3 / MPO / both) OR (biopsy = Definite vasculitis AND histologically confirmed = YES)


```{r, warning=FALSE}

Inclusion_Criteria_rkd <- DemographicFilterRKD(clean_rkd, paste(a,"/output",sep=""))
head(Inclusion_Criteria_rkd)[1:5,1:10]
```

## Classify the RKD patient

The ClassifyRKDCases function use different algorithm create by several previous employee:

* "BRelapse" algorithm perform the prediction of a Relapse. The output of the function is a variable that tell if the patient have a definite relapse, a possible relapse or no relapse fr one encounter.
- 

```{r, warning=FALSE}

Classify_rkd <- ClassifyRKDCases(clean_rkd, paste(a,"/output",sep=""), "BRelapse")
head(Classify_rkd)[1:5,1:10]
```

## Filter with the classification

The ClinicalFilterRKD function is a filter based on classify algorithm

```{r, warning=FALSE}

Filter_Classify_rkd <- ClinicalFilterRKD(Classify_rkd, paste(a,"/output",sep=""), "BRelapse")
head(Classify_rkd)[1:5,1:10]
```


## Load Freezer Work data

This function load the data from Freezerwork database based on their blood matrix:

* Plasma
- PBMC
- Serum
- DNA
- urine
- RNA

```{r, warning=FALSE}

Load_FW <- LoadFW(fw_path, "Plasma")
head(Load_FW)[1:5,1:10]
```

## Cleanning Freezer Work data

The CleanFW function is done in 2 step:

* Put the date in the same format of the RKD data
- Remove the sample that no amount
- Remove sample without RKD.ID

```{r, warning=FALSE}

Clean_FW <- CleanFW(Load_FW, paste(a,"/output",sep=""))
head(Clean_FW)[1:5,1:10]
```

## Finding the good sample

The Goal of the FindFWSample function is to find the best sample to measure to answer to the statistician question. It is just a merge between the 2 files.

```{r, warning=FALSE}

Finding_Good_Sample <- FindFWSample(Clean_FW, Inclusion_Criteria_rkd , paste(a,"/output",sep=""))
head(Finding_Good_Sample)[1:5,1:10]
```

# Appendix

```{r sessioin}
sessionInfo()
```